---
title: "Test correlation of IHC measurements with cluster signatures in IPF core samples"
author: "[Alex Abbas](mailto:abbas@gene.com)"
date: "`r Sys.Date()`"
output:
    html_document:
      toc: yes
      toc_depth: 4
      theme: flatly
      css : http://research.gene.com/people/abbas/css/tocWide.css
---

```{r global_options, include=F, echo=F}
knitr::opts_chunk$set(fig.width=6, fig.height=6, fig.path='Figs/', echo=F, warning=FALSE, message=FALSE, include=T)
```

```{r load_packages, include=F, echo=F}
library(DT)
library(knitr)
library(ggplot2)
library(Biobase)
library(reshape)
library(GGally)
library(ellipse)
library(RColorBrewer)

```

```{r package_options, include=F, echo=F}
knitr::opts_knit$set(progress = TRUE, verbose = TRUE)
# library(rmarkdown)
# library(BiocStyle)
# render('corIhcClust.Rmd', "html_document", '../results/corIhcClust.html')
```

```{r etc, include=F, echo=F}
calcSignatureScores = function(dat=v,geneids=rownames(dat)) {
  if (length(geneids)==1) {
   return(dat[geneids,])
  }
  sliceids = geneids[geneids %in% rownames(dat)]
  sigdat = dat[sliceids,]
  pc1 = prcomp(t(sigdat))$x[,1]
  meansig = apply(sigdat,2,mean)
  if ( cor(pc1,meansig) < 0 ) { pc1 = - pc1 }
  pc1 = pc1 * sd(meansig)/sd(pc1)
  pc1 = pc1 + mean(meansig) - mean(pc1)
  names(pc1) = colnames(dat)
  pc1
}
myminor = unlist(lapply(c(-3:5),function(o){c(2:9)*10^o}))
```

Hypothesis
===
Previous IPF biopsies and lung sections have identified expression signatures related to diagnosis or histology. Test whether various expression signatures are related to cellular populations quantified by IHC.

Analysis
========
```{r loadData, echo=F, include=F}
load("../data/eset.RData")
pheno = pData(eset)
exprdat = exprs(eset)
```

Calculate signatures
-------------
Examine distribution of k-means cluster signatures by diagnosis
```{r sigs, fig.width=8, fig.height=8, echo=F}
source("../R/functions.R")
k8 = readDelim("../data/geneClusterAnnot.txt")
k8genesets = lapply(c(1:8),function(i){as.character(subset(k8,Cluster==i)$ID)})
k8NameDat = readDelim("../data/k8names.txt")
clustnames = paste(c(1:8),k8NameDat$name)
names(k8genesets) = clustnames
k8Scores = data.frame(
 SAMID = colnames(exprdat),
 diagnosis = pheno$TISSUE_DIAGNOSIS,
 do.call(cbind,lapply(k8genesets,function(asig){calcSignatureScores(exprdat,asig)}))
)
colnames(k8Scores)[3:10] = paste(gsub(" ","_",k8NameDat$name),k8NameDat$i,sep="_")
k8long = melt(k8Scores, id.vars=c('SAMID','diagnosis'), value.name='signature')
colnames(k8long)[c(3:4)] = c("signature","score")
ggplot(k8long, aes(diagnosis, score)) + 
 stat_summary(fun.y = mean, geom='point') + 
 stat_summary(fun.data = 'mean_cl_boot', geom='errorbar', width=.25) + 
 facet_wrap(~signature,nrow=2) + scale_y_log10(minor_breaks = myminor)
```

IHC pairs plot
-------------
Examine intercorrelation of IHC variables
```{r ihcpairs, fig.width=8, fig.height=8, echo=F}
ihc = pheno[,c("Neutrophil","Macrophage_CD68","Bcell_CD79a","CD4","CD8","Eosinophils")]
ggpairs(ihc, columns=c("Neutrophil","Macrophage_CD68","Bcell_CD79a","CD4","CD8","Eosinophils"),
    diag=list(continuous="densityDiag", discrete="barDiag"), axisLabels="show")
```

IHC vs. signatures overall
-------------
Correlate IHC against signature scores
```{r ihcSig, fig.width=8, fig.height=8, echo=F}
k8ihc = cbind(k8Scores,ihc)

my_colors <- brewer.pal(5, "Spectral")
my_colors=colorRampPalette(my_colors)(100)
 
# Order the correlation matrix
data = cor(k8ihc[,3:16])
ord <- order(data[1, ])
data_ord = data[ord, ord]
plotcorr(data_ord , col=my_colors[data_ord*50+50] , mar=c(1,1,1,1)  )
```

IHC vs. signatures within IPF
-------------
Correlate IHC against signature scores using only data from IPF patients
```{r ihcSigIpf, fig.width=8, fig.height=8, echo=F}
k8ihc = cbind(k8Scores,ihc)
k8ihcIpf = subset(k8ihc,diagnosis == "IPF")

data = cor(k8ihcIpf[,3:16])
ord <- order(data[1, ])
data_ord = data[ord, ord]
plotcorr(data_ord , col=my_colors[data_ord*50+50] , mar=c(1,1,1,1)  )
```

Lymphocyte sig and B cells
-------------
The CD79 count and the lymphocyte signature are correlated at r=0.54 within IPF samples.

```{r cd79Th2, fig.width=8, fig.height=7, echo=F}
ggplot(k8ihc, aes(x=Bcell_CD79a, y=lymphocyte_4, color=diagnosis)) + geom_point(shape=1) + geom_smooth(method=lm) +
 scale_x_log10(minor_breaks = myminor) + scale_y_log10(minor_breaks = myminor)
```


Modules & Info
==============
```{r furlaniModules}
system("module list 2>&1",intern=T)
```

```{r sessioninfo}
sessionInfo()
```




